Graphical Tests for Power Comparison of Competing Designs

### 2012-03-28 setup.tex
results for experiment 1 done

XXX picture of sampling size?
An additional factor we were interested in this particular situation was to assess, how much of the data we actually needed to use in a design to have observers pick out the pattern. Clearly, a design is more efficient, if a smaller sample size is sufficient for showing the presence of a relationship. In order to investigate this, we took different size samples of the data and created lineup plots of all four designs for these subsets. Another perturbation to the data are `shifts' in wind direction, i.e. we make use of the circular nature of the wind direction and adjust the $x$ axis by using different offsets.
XXXX show picture of offsets
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{euclid-offset} 
   \includegraphics[width=\linewidth]{polar-offset} 
   \caption{ \label{fig:offset} (Top) Euclidean charts of the same sample, with shifts in wind direction by 90, 180, and 270 degrees. The shift by 90 and 270 leads to a `valley' versus a `mountain' pattern, whereas a shift by 180 degrees inverts the wave pattern from `down-up' to `up-down'.\hfill\newline
      (Bottom) Shifts in wind direction in the polar chart lead to rotations of the display by 90 degrees.}
\end{figure}

 This results in a shift of the wave in Euclidean charts and a rotation in the polar charts. Our initial hypothesis was that this would have no effect on the polar charts, but might have a deteriorating effect on the Euclidean charts, in the case that the peak or the valley of the wave was at either end of the $x$ axis. 
While these simple perturbations and different sample sizes allow us to get insight into different aspects of the designs, they also allow us to get multiple responses from each observer to assess individual ability without the need to go outside the framework of the original data, i.e. we leave the joint relationship between $x$ and $y$ essentially unchanged. For all of these combinations we produced two replicates, resulting in a total of 192 different lineups ($ 4 \text{ designs} \times 6 \text{ sample sizes} \times 4 \text{ offsets} \times 2 \text{ reps})$.

@@

XXX confidence by test, fill by correct
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{conf-answer} 
   \caption{Barcharts of (subjective) confidence in answering correctly for each lineup. A strongly bimodal distribution is apparent, while confidence levels for polar charts are more uniform. There are only slight (and non significant) difference between confidence levels and correctness of results.}
   \label{fig:time}
\end{figure}

@@

Time taken to answer was log transformed before the modelling process to de-emphasize the impact of very large values (up to 500 sec). The findings are consistent with correctness - time taken shows big differences between polar and Euclidean charts. The reference lines seem to increase the evaluation time, but not significantly. An increase in sample size decreases evaluation time, The inversion of the wave pattern at 180 degrees leads to a significant increase in time for Euclidean charts, but not for polar charts (at least not significantly).

@@

>>>>

An additional factor we were interested in this particular situation was to assess, how much of the data we actually needed to use in a design to have observers pick out the pattern. Clearly, a design is more efficient, if a smaller sample size is sufficient for showing the presence of a relationship. In order to investigate this, we took different size samples of the data and created lineup plots of all four designs for these subsets. The effect of sampling size on the displays mostly stems from the additional variability introduced when using small samples, which might hide the pattern, if it is not displayed prominently. 
XXX picture of sampling size?
Another perturbation to the data are `shifts' in wind direction, i.e. we make use of the circular nature of the wind direction and adjust the $x$ axis by using different offsets.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{euclid-offset} 
   \includegraphics[width=\linewidth]{polar-offset} 
   \caption{ \label{fig:offset} (Top) Euclidean charts of the same sample, with shifts in wind direction by 90, 180, and 270 degrees. The shift by 90 and 270 leads to a `valley' versus a `mountain' pattern, whereas a shift by 180 degrees inverts the wave pattern from `down-up' to `up-down'.\hfill\newline
      (Bottom) Shifts in wind direction in the polar chart lead to rotations of the display by 90 degrees.}
\end{figure}

 This results in a shift of the wave in Euclidean charts and a rotation in the polar charts as can be seen in figure \ref{fig:offset}. Our initial hypothesis was that this would have no effect on the polar charts, but might have a deteriorating effect on the Euclidean charts, in the case that the peak or the valley of the wave was at either end of the $x$ axis. 
While these simple perturbations and different sample sizes allow us to get insight into different aspects of the designs, they also allow us to get multiple responses from each observer to assess individual ability without the need to go outside the framework of the original data, i.e. we leave the joint relationship between $x$ and $y$ essentially unchanged. For all of these combinations we produced two replicates, resulting in a total of 192 different lineups ($ 4 \text{ designs} \times 6 \text{ sample sizes} \times 4 \text{ offsets} \times 2 \text{ reps})$.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{conf-answer} 
   \caption{Barcharts of (subjective) confidence in answering correctly for each lineup. A strongly bimodal distribution is apparent, while confidence levels for polar charts are more uniform. There are only slight (and non significant) difference between confidence levels and correctness of results.}
   \label{fig:time}
\end{figure}

@@

Time taken to answer was log transformed before the modeling process to de-emphasize the impact of very large values (up to 500 sec). The findings are consistent with correctness - time taken shows big differences between polar and Euclidean charts. The reference lines seem to increase the evaluation time, but not significantly. An increase in sample size decreases evaluation time, The inversion of the wave pattern at 180 degrees leads to a significant increase in time for Euclidean charts, but not for polar charts (at least not significantly).

@@

\begin{figure*}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{predict-power} 
   \caption{Predicted Power of designs.  The thin lines and the points on $y$ axis show  variability due to individuals' abilities. The saturated lines show average predicted power for each of the designs.}
   \label{fig:power}
\end{figure*}
